---
title: "Using ThermStats to calculate statistics for thermal images"
output: rmarkdown::html_vignette
author: Rebecca A. Senior
date: "`r format(Sys.time(), '%d %B, %Y')`"
vignette: >
  %\VignetteIndexEntry{Using ThermStats to calculate statistics for thermal images}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# tl;dr

`ThermStats` is designed for biologists using thermography to quantify thermal heteorgeneity. It uses the `Thermimage` package [@tattersall_thermimage:_2017] to batch process data from FLIR thermal cameras, and takes inspiration from FRAGSTATS, `SDMTools`, @faye_toolbox_2016 and @shi_framework_2016 to facilitate the calculation of various metrics of thermal heteroeneity for any gridded temperature data.

The package is available to download from GitHub using `devtools`:

```{r install}
devtools::install_github("rasenior/ThermStats")
library(ThermStats)
```

# Extracting raw data

# Converting raw data to temperature

# Calculating thermal statistics

# Plotting









